---
title: "count_md"
author: "Daphne Chevalier"
date: "2024-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Packages
```{r}
library(ggplot2)
library(ggpubr)
library(ggsignif)
library(data.table)
library(igraph)
library(scales)
library(vegan)
library(RColorBrewer)
library(lme4)
library(DHARMa)
library(performance)
library(mvabund)
library(emmeans)
library(glmmTMB)
library(dlookr)
library(multcomp)
library(multcompView)
library(stringr)
```

## Setup

### Count data download

```{r}
all_pitfall_names <- fread("../raw_data/pitfall_names_241024.csv", header = FALSE)

setnames(all_pitfall_names, "V1", "FullName")

all_pitfall_names <- all_pitfall_names[grep("debris", FullName, invert = TRUE)]
all_pitfall_names <- all_pitfall_names[grep("multiple_insects", FullName, invert = TRUE)]
all_pitfall_names$FullName <- gsub(" ", "_", all_pitfall_names$FullName)
all_pitfall_names[, c("Root","Filename") := tstrsplit(FullName, "\\/+(?!\\S*\\/)", perl = TRUE)]
all_pitfall_names[, c("Root","Phylum","Class","Order","Family","Subfamily","Genus","Species") := tstrsplit(Root, "/")]
all_pitfall_names[, c("StartDate", "Site", "PrescribedTreatment", "Distance") := tstrsplit(Filename, "_", fixed = TRUE, type.convert = TRUE)[1:4]]
all_pitfall_names[, Distance := strtrim(Distance, 2)]
all_pitfall_names[, StartDate := as.Date(StartDate)]
all_pitfall_names[, id := .I]
all_pitfall_names <- all_pitfall_names[StartDate > as.Date("2023-06-21")]

metatable <- fread("../cleaned_data/metadata.csv")
metatable <- metatable[!is.na(Site)]

pitfall_names <- all_pitfall_names[metatable, Treatment := i.Treatment, on=.(Site, StartDate)]
pitfall_names <- pitfall_names[Treatment %in% c("A","C","W")]
pitfall_names$Treatment <- factor(pitfall_names$Treatment, levels=c("C","A","W"), ordered = FALSE)

pitfall_names <- pitfall_names[Class != "unclassified_arthropod"]
```

### Biomass data download
```{r}
biomass_c <- fread("../cleaned_data/biomass.csv")
biomass_c <- biomass_c[, StartDate := as.Date(StartDate)]
setnames(biomass_c, "Near", "0m")
setnames(biomass_c, "Far", "5m")
biomass_c <- melt(setDT(biomass_c[, !c("Total")]), id.vars = c("StartDate","Site", "Treatment"), variable.name = "Distance")
setnames(biomass_c, "value", "Biomass")

first_julian_biomass <- julian(min(biomass_c$StartDate))
biomass_c[, StartDateJulian := as.numeric(julian(StartDate)) - as.numeric(first_julian_biomass)]
biomass_c[,StartDateJulian := scale(StartDateJulian)]
biomass_c

biomass_zeroes <- biomass_c[Biomass == 0]
biomass_zeroes[, Phylum := "Arthropod"]
setnames(biomass_zeroes, "Biomass", "Count")

merge(pitfall_names, biomass_zeroes, all.x = TRUE, all.y = TRUE)

biomass_c$Treatment <- factor(biomass_c$Treatment, levels=c("C","W","A"), ordered = FALSE)

taxa_cols <- c("Class","Order","Family","Subfamily","Genus","Species")

my_comparisons <- list(c("C","W"), c("C","A"), c("W","A"))
```

### Add rows for each nested level
We needed our table to be exhaustive, with a row for every taxonomic group. For example, if we counted x individuals of Species S within Genus G within Family F in a given pitfall trap, we wanted a row for each of these groups.
```{r}
wide_counts <- pitfall_names[, !c("FullName", "Root", "Filename", "PrescribedTreatment", "id")][, .N, by=.(Phylum, Class, Order, Family, Subfamily, Genus, Species, Site, Treatment, Distance, StartDate)]
cols <- c("Phylum", "Class", "Order", "Family", "Subfamily", "Genus", "Species")
f <- function(dt) {
  i <- sum(!is.na(dt[,..cols]))
  if (i > 1) {
    dt <- dt[rep(1L, i)]
    for (j in 2:i) set(dt, i = 1:(j - 1), j = j, NA)
  }
  dt
}
wide_counts <- wide_counts[,f(.SD), 1:nrow(wide_counts)][,nrow := NULL][
  ,.(N = sum(N)), Phylum:StartDate
]

setnames(wide_counts, old = "N", new = "Count")
```

### Add Julian date and order treatments
```{r}
first_julian <- julian(min(wide_counts$StartDate))
wide_counts[, StartDateJulian := as.numeric(julian(StartDate)) - as.numeric(first_julian)]
wide_counts[,StartDateJulian := scale(StartDateJulian)]
wide_counts

wide_counts$Treatment <- factor(wide_counts$Treatment, levels=c("C","W","A"), ordered = FALSE)

```

## Size calculation

```{r}
capture_size <- wide_counts[Phylum == "Arthropoda" & is.na(Class)]
capture_size <- capture_size[biomass_c, on = .(StartDate, Site, Treatment, Distance)]
capture_size <- capture_size[Phylum == "Arthropoda"]

capture_size[, Size := Biomass/Count * 1000]

capture_size$Treatment <- factor(capture_size$Treatment, levels=c("C","W","A"), ordered = FALSE)
```

## Boxplots
### Old
```{r}
# Count
pdf("../plots/count_box_all.pdf", w = 15, h = 9)
ggplot(capture_size, aes(Treatment,Count)) +
  geom_boxplot(outlier.shape = NA) +
  theme(axis.text=element_text(size=24),
        axis.title=element_text(size=24,face="bold"),
        title = element_text(size=25, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "More arthropods captured under white light than in darkness") +
  stat_compare_means(method = "wilcox.test", comparisons = my_comparisons, size = 8)

# Size
pdf("../plots/size_box_all.pdf", w = 15, h = 9)
ggplot(capture_size, aes(Treatment,Size)) +
  geom_boxplot(outlier.shape = NA) +
  theme(axis.text=element_text(size=22),
        axis.title=element_text(size=22,face="bold"),
        title = element_text(size=24, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Larger arthropods (on average) captured under lit conditions") +
  ylab("Size (mg/arthropod)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_compare_means(method = "wilcox.test", comparisons = my_comparisons, size = 8)
dev.off()
```

### Setup for faceted biomass/count/size boxplots
```{r}
bsc <- melt(capture_size[, c("Site", "Treatment", "Distance", "StartDateJulian", "Biomass", "Count", "Size")], 
     id.vars = c("Site", "Treatment", "Distance", "StartDateJulian"), variable.name = "Measure")

box_labels <- c('Biomass' = 'Arthropod biomass (g)', 'Count' = 'Total arthropod count', 'Size' = 'Arthropod biomass/count (mg)')
```

### Biomass, count, and size
```{r}
pdf("../plots/bsc_box_all.pdf", w = 15, h = 9)
ggplot(bsc, aes(Treatment, value)) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = position_jitter(seed = 1, width = 0.2)) +
  theme(panel.grid = element_blank()) +
  labs(y="") +
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=22,face="bold"),
        title = element_text(size=22, face = "bold"),
        strip.text = element_text(size = 13)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans='log10') +
    scale_x_discrete(labels = c("No light", "White", "Amber")) +
  #stat_compare_means(method = "wilcox.test", comparisons = my_comparisons) +
  facet_grid(rows = vars(Measure), scales = "free_y", labeller = labeller(Measure = box_labels))
dev.off()
```

### Count and size
```{r}
pdf("../plots/countsize_box_all.pdf", w = 15, h = 9)
ggplot(bsc[Measure == "Count" | Measure == "Size"], aes(Treatment, value)) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = position_jitter(seed = 1, width = 0.2)) +
  theme(panel.grid = element_blank()) +
  labs(y="") +
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=22,face="bold"),
        title = element_text(size=22, face = "bold"),
        strip.text = element_text(size = 13)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans='log10') +
  scale_x_discrete(labels = c("No light", "White", "Amber")) +
  #stat_compare_means(method = "wilcox.test", comparisons = my_comparisons) +
  facet_grid(rows = vars(Measure), scales = "free_y", labeller = labeller(Measure = box_labels))
dev.off()
```

### Biomass
```{r}
pdf("../plots/biomass_box_all.pdf", w = 15, h = 9)
ggplot(bsc[Measure == "Biomass"], aes(Treatment, value)) +
  geom_boxplot() +
  #geom_point(alpha = 0.3, position = position_jitter(seed = 1, width = 0.2)) +
  theme(panel.grid = element_blank()) +
  labs(y = "Biomass (grams)", title = "Biomass captured across treatments") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=22,face="bold"),
        title = element_text(size=22, face = "bold"),
        strip.text = element_text(size = 13)) +
  ylim(0, 15) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels = c("No light", "White", "Amber"))
dev.off()
```

### Count
```{r}
pdf("../plots/count_box_all.pdf", w = 15, h = 9)
ggplot(capture_size, aes(Treatment,Count)) +
  geom_boxplot(outlier.shape = NA) +
  theme(axis.text=element_text(size=24),
        axis.title=element_text(size=24,face="bold"),
        title = element_text(size=25, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "More arthropods captured under white light than in darkness") +
  stat_compare_means(method = "wilcox.test", comparisons = my_comparisons, size = 8)
dev.off()
```

### Classes
```{r}
pdf("../plots/classes.pdf", w = 15, h = 9)
ggplot(wide_counts[!is.na(Class) & is.na(Order), sum(Count), by=Class], aes(x = Class, fill = Class, y = V1)) +
  geom_col() +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(axis.text=element_text(size=24),
        axis.title=element_text(size=24,face="bold"),
        title = element_text(size=25, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none") +
  scale_y_continuous(trans='log10') +
  labs(title = "Number of arthropods captured across classes") +
  ylab("Count (log scale)")
dev.off()
```
### Insect orders
```{r}
pdf("../plots/orders.pdf", w = 15, h = 9)
ggplot(wide_counts[Class == "Insecta" & !is.na(Order) & is.na(Family), sum(Count), by=Order], aes(x = Order, fill = Order, y = V1)) +
  geom_col() +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(axis.text=element_text(size=24),
        axis.title=element_text(size=24,face="bold"),
        title = element_text(size=25, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none") +
  scale_y_continuous(trans='log10') +
  labs(title = "Number of insects captured across orders") +
  ylab("Count (log scale)")
dev.off()
```
## Simple biomass model
```{r}
#lme_biomass <- glmmTMB(Biomass ~ Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
lme_biomass <- glmmTMB(sqrt(Biomass) ~ Treatment * Distance,
                      data = biomass_c)
plot(simulateResiduals(lme_biomass))
summary(lme_biomass)
emmgrid <- emmeans(lme_biomass, specs=pairwise~Treatment*Distance, type="response")
emmgrid$contrasts
emmeans(lme_biomass, list(pairwise ~ Treatment), adjust = "tukey")
```


## Pooled count model, only treatment
```{r}
lme_count_lim <- glmmTMB(Count ~ Treatment,
                      data = wide_counts[Phylum == "Arthropoda" & is.na(Class)],
                     family = "nbinom2")
plot(simulateResiduals(lme_count_lim))
#r2_nakagawa(lme_count)
summary(lme_count_lim)
emmeans(lme_count_lim, list(pairwise ~ Treatment), adjust = "tukey")
```

## Pooled count model
```{r}
lme_count <- glmmTMB(Count ~ Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                      data = wide_counts[Phylum == "Arthropoda" & is.na(Class)],
                     family = "nbinom2")
plot(simulateResiduals(lme_count))

#r2_nakagawa(lme_count)
summary(lme_count)
emmeans(lme_count, list(pairwise ~ Treatment), adjust = "tukey")
```

## Pooled size model
```{r}
lme_size <- lmer(sqrt(Size) ~ Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance), data = capture_size)
summary(lme_size)
plot(simulateResiduals(lme_size))
emmeans(lme_size, list(pairwise ~ Treatment), adjust = "tukey")
anova(lme_size)
r2_nakagawa(lme_size)
```

# Taxon-specific models
## Classes
Add zero rows.
```{r}
zero_classes <- wide_counts[!is.na(Class) & is.na(Order) & !is.na(Treatment), .SD, .SDcols = c("Class", "Site", "Treatment", "Distance", "StartDateJulian", "Count")]

zero_classes <- zero_classes[
  CJ(Class = Class,
     Site = Site,
     Distance = Distance,
     StartDateJulian = StartDateJulian,
     unique = TRUE), 
  on = .(Class, Site, Distance, StartDateJulian)
]

zero_classes[, .N, by=.(Site, StartDateJulian, Class, Distance)]

setnafill(zero_classes, fill = 0, cols=c("Count"))
```





Select most-frequently-occurring taxa.
```{r}
classes <- zero_classes[, sum(Count), by=Class][zero_classes, on=c("Class")][V1>0.03*sum(unique(V1))][,.SD, .SDcols = !c('V1')]
factor_cols <- c("Class", "Distance", "Site")
classes[, (factor_cols) := lapply(.SD, factor, ordered = FALSE), .SDcols = factor_cols]
```

Model.
```{r}
lme_class_1 <- glmmTMB(Count ~ Class + Treatment : Class + Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                     data = classes,
                     ziformula= ~1,
                     family="nbinom2")
plot(simulateResiduals(lme_class_1))
testZeroInflation(simulateResiduals(lme_class_1))
testDispersion(simulateResiduals(lme_class_1), type = "DHARMa")
testOutliers(simulateResiduals(lme_class_1))

lme_class <- glmmTMB(Count ~ Class * Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                     data = classes,
                     ziformula= ~1,
                     family="nbinom2")
plot(simulateResiduals(lme_class))
```

Plot.
```{r}
emmeans(lme_class_1, ~Treatment * Class, type="response")
plot(emmeans(lme_class_1, ~Treatment | Class, type="response"), comparisons = TRUE)
pairs(emmeans(lme_class_1, ~Class * Treatment, type="response"), simple = "each", adjust = "holm")
```



Per trap.
```{r}
# we want a stacked bar graph
classes[, .N, by=c("StartDateJulian", "Treatment")]
classes[Count > 0, .N, by=c("Treatment", "Class")]
classes[, .(Traps = .N), by=.(Class, Distance, Treatment)]
#[, .(Number = sum(Count), PerTrap = sum(Count)/Traps), by=.(Class, Distance, Treatment)]
pitfall_classes <- classes[, .(Traps = .N, Number = sum(Count), PerTrap = sum(Count)/.N), by=.(Class, Distance, Treatment)]

```


## Families
Get the families, add zero rows.
```{r}
families <- wide_counts[!is.na(Family) & is.na(Subfamily), .SD, .SDcols = c("Family", "Site", "Treatment", "Distance", "StartDateJulian", "Count")]
families <- families[
  CJ(Family = Family,
     Site = Site,
     Distance = Distance,
     StartDateJulian = StartDateJulian,
     unique = TRUE), 
  on = .(Family, Site, Distance, StartDateJulian)
]
setnafill(families, fill = 0, cols=c("Count"))
```

Filter on frequency.
```{r}
families_freq <- families[families[, sum(Count), by=Family][V1>0.01*sum(unique(V1))], on=c("Family")]
```

Model.
```{r}
lme_fam_1 <- glmmTMB(Count ~ Family + Treatment : Family + Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                     data = families_freq,
                     ziformula=~1,
                     family=nbinom2)

plot(simulateResiduals(lme_fam_1))
```

Plot.
```{r}
plot(emmeans(lme_fam_1, ~Treatment | Family, type="response"), comparisons = TRUE)
pairs(emmeans(lme_fam_1, ~Family * Treatment, type="response"), simple = "each", adjust = "holm")
```

Other model.
```{r}
lme_fam_dist <- glmmTMB(Count ~ Family * Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                         data = families_freq,
                         ziformula=~1,
                         family=nbinom2)
summary(lme_fam_dist)
plot(emmeans(lme_fam_dist, ~Treatment | Family, type="response"), comparisons = TRUE)

pairs(emmeans(lme_fam_dist, ~Family * Treatment, type="response"), simple = "each", adjust = "holm")
fam_dist_emms <-  emmeans(lme_fam_dist, ~ Treatment * Distance | Family)
contrast(fam_dist_emms, interaction = "pairwise")
```

## Subfamilies
Get the subfamilies, add zero rows.
```{r}
subfamilies <- wide_counts[!is.na(Subfamily) & is.na(Genus), .SD, .SDcols = c("Subfamily", "Site", "Treatment", "Distance", "StartDateJulian", "Count")]
subfamilies <- subfamilies[
  CJ(Subfamily = Subfamily,
     Site = Site,
     Treatment = Treatment,
     Distance = Distance,
     StartDateJulian = StartDateJulian,
     unique = TRUE), 
  on = .(Subfamily, Site, Treatment, Distance, StartDateJulian)
]

setnafill(subfamilies, fill = 0, cols=c("Count"))
```

Filter for frequency.
```{r}
subfamilies_freq <- subfamilies[subfamilies[, sum(Count), by=Subfamily][V1>0.01*sum(unique(V1))], on=c("Subfamily")]
```

Model.
```{r}
lme_sub_1 <- glmmTMB(Count ~ Subfamily * Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
#lme_sub_1 <- glmmTMB(Count ~ Subfamily + Treatment : Subfamily + Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                     data = subfamilies_freq,
                     ziformula=~1,
                     family=nbinom2)
#diagnose(lme_sub_1)
plot(simulateResiduals(lme_sub_1))
# testZeroInflation(simulateResiduals(lme_sub_1))
# testDispersion(simulateResiduals(lme_sub_1), type = "DHARMa")
# testOutliers(simulateResiduals(lme_sub_1))
```

Plot.
```{r}
emmeans(lme_sub_1, ~Treatment * Subfamily, type="response")
pairs(emmeans(lme_sub_1, ~Subfamily * Treatment, type="response"), simple = "each", adjust = "holm")

plot(emmeans(lme_sub_1, ~Treatment | Subfamily, type="response"), comparisons = TRUE)
```


## Genera
Get the genera, add zero rows.
```{r}
genera <- wide_counts[!is.na(Genus) & is.na(Species), .SD, .SDcols = c("Genus", "Site", "Treatment", "Distance", "StartDateJulian", "Count")]
genera <- genera[
  CJ(Genus = Genus,
     Site = Site,
     Treatment = Treatment,
     Distance = Distance,
     StartDateJulian = StartDateJulian,
     unique = TRUE), 
  on = .(Genus, Site, Treatment, Distance, StartDateJulian)
]

genera <- setnafill(genera, fill = 0, cols=c("Count"))
```

Filter for frequency.
```{r}
genera_freq <- genera[genera[, sum(Count), by=Genus][V1>0.03*sum(unique(V1))], on=c("Genus")]
```

Model.
```{r}
lme_gen_1 <- glmmTMB(Count ~ Genus * Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
#lme_gen_1 <- glmmTMB(Count ~ Genus + Treatment : Genus + Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                     data = genera_freq,
                     ziformula=~1,
                     family=nbinom2)
summary(lme_gen_1)
plot(simulateResiduals(lme_gen_1))
testZeroInflation(simulateResiduals(lme_gen_1))
testDispersion(simulateResiduals(lme_gen_1), type = "DHARMa")
testOutliers(simulateResiduals(lme_gen_1))
```

Plot.
```{r}
emmeans(lme_gen_1, ~Treatment * Genus, type="response")
pairs(emmeans(lme_gen_1, ~Genus * Treatment, type="response"), simple = "each", adjust = "holm")

plot(emmeans(lme_gen_1, ~Treatment | Genus, type="response"), comparisons = TRUE)
```




## species
Get the species, add zero rows.
```{r}
species <- wide_counts[!is.na(Species), .SD, .SDcols = c("Species", "Site", "Treatment", "Distance", "StartDateJulian", "Count")]
species <- species[
  CJ(Species = Species,
     Site = Site,
     Treatment = Treatment,
     Distance = Distance,
     StartDateJulian = StartDateJulian,
     unique = TRUE), 
  on = .(Species, Site, Treatment, Distance, StartDateJulian)
]

species <- setnafill(species, fill = 0, cols=c("Count"))
```

Filter for frequency.
```{r}
species_freq <- species[species[, sum(Count), by=Species][V1>0.03*sum(unique(V1))], on=c("Species")]
```

Model.
```{r}
lme_spec_1 <- glmmTMB(Count ~ Species + Treatment : Species + Treatment * Distance + StartDateJulian + (1|Site) + (1|Site:Distance),
                     data = species_freq,
                     ziformula=~1,
                     family=nbinom2)
summary(lme_spec_1)
plot(simulateResiduals(lme_spec_1))
testZeroInflation(simulateResiduals(lme_spec_1))
testDispersion(simulateResiduals(lme_spec_1), type = "DHARMa")
testOutliers(simulateResiduals(lme_spec_1))
```

Plot.
```{r}
emmeans(lme_spec_1, ~Treatment * Species, type="response")
pairs(emmeans(lme_spec_1, ~Species * Treatment, type="response"), simple = "each", adjust = "holm")

plot(emmeans(lme_spec_1, ~Treatment | Species, type="response"), comparisons = TRUE)
```



## Plotting emmeans
```{r}
lme_fam_1_summary <- summary(emmeans(lme_fam_1, ~Treatment : Family, type="response"), 
                             comparisons = TRUE)

lme_fam_1_means <- emmeans(object = lme_fam_1,
                             specs = ~ Treatment : Family,
                           type = "response")

lme_fam_1_cld <- cld(object = lme_fam_1_means,
                       adjust = "Tukey",
                       Letters = letters,
                       alpha = 0.05)
lme_fam_1_cld
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
